\chapter{Introduction}

\newthought{Sensible Calculus} aims to satisfy the questioner and the skeptic. There are many aspects of Calculus and a primary focus for many textbooks are a mixture of formulaic solving and some attempt to use this as a practical tool. 

But Calculus is a fascinating journey into infinity, rooted in the real world. Our interactions with the world, our knowledge of the world, appears to be finite. But by working with infinity, there are many impossibly seeming tasks that we can get a handle on. Since we are finite, infinity is never realized. It is therefore important to be clear exactly what we are talking about. 

This books starts with a review of arithmetic up through rational numbers. While this material can be skipped, it can be useful to appreciate the foundations for all the concrete numbers that we will use. It also reviews the algebra and geometry which can be done with rational numbers alone. 

The most novel part of this book is Part II which is delving into real numbers. The approach taken here is that of representing real numbers by the intervals that contain them. 

Part III explores functions from a definitional point of view. Conitnuity will be front and center here. 

Part IV explores calculus itself.  

Part V explores multivariable considerations. This is just a dabbling. 

Part VI discuss some higher level concerns, a smattering of topics from analysis and numerical analysis that might be of some interest. 

The Appendices contain some interesting, but tangential material. The first appendix is the construction of the natural numbers from set theory. While the main approach of using counting is fine, this approach may appeal to those wishing to minimize starting assumptions. Another appendix concerns the construction of the RatMath engine. 